{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}

### 2016-12-11 sftrends_manu.Rnw
added seas by year to trend map

source('R/funcs.r')
library(WRTDStidal)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(grid)
library(purrr)
library(tidyr)
library(Hmisc)
library(ggrepel)
library(ggtree)
@

@@

<<trndmap, fig.cap = 'Percent change per year in nitrogen analytes for (a) annual and (b) seasonal aggregations.  Changes are based on seasonal Kendall tests of flow-normalized results within each time period. Station names are shown in the top left panel. Three values less than 5\\% change were truncated for visual clarity of the bar plots, see \\cref{tab:trnds} for actual values. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.', fig.height = 8, fig.width = 7, cache = T, out.width = '\\textwidth'>>=

res = c('din', 'nh', 'no23')
mobrks <- list(c(3, 4, 5), c(6, 7, 8), c(9, 10, 11), c(12, 1, 2))
yrbrks <- c(-Inf, 1995, Inf)
molabs <- c('Spring', 'Summer', 'Fall', 'Winter')
yrlabs <- c('1976-1995', '1996-2014')
buffx = 0.00035
buffy = 0.0022
cols =  c('#D53E4F', '#4DAF4A')
sz_rng = c(2, 11)
shps = c(24, 25)
ndivs = 6
strp_fl = 'grey90'
bar = FALSE
stat_labs = TRUE
leg = TRUE
leg_lab = 'Percent change'

# load required data
data(delt_dat)
data(delt_map)
data(trnds_sk)

@@

# bounding box
lims<- select(statmeta, -Site_Code) %>% 
  apply(., 2, function(x) range(x, na.rm = TRUE))
lims <- c(
  (1 + buffx) * lims[1, 2], 
  (1 - buffy) * lims[1, 1], 
  (1 - buffx) * lims[2, 2], 
  (1 + buffy) * lims[2, 1]
  )    
names(lims) <- c('left', 'bottom', 'right', 'top')

# response labels 
reslabs <- c('din an', 'nh an', 'no23 an', 'din mo', 'nh mo', 'no23 mo')
reslevs <- c(
  'DIN', 
  expression(paste(NH [4] ^ '+')),
  expression(paste(NO [2] ^ '-', ' / ', NO [3] ^ '2-')),
  'DIN', 
  expression(paste(NH [4] ^ '+')),
  expression(paste(NO [2] ^ '-', ' / ', NO [3] ^ '2-'))
)

@@

yrng <- c(-5, 5)

@@

# unnest data frame, assign colors by rank and sign
toplo <- select(trnds_sk, Site_Code, resvar, resvar, cat, ztest, perchg) %>% 
  left_join(., statmeta, by = 'Site_Code') %>% 
  data.frame %>% 
  mutate(
    perchg = pmax(yrng[1], perchg),
    perchg = pmin(yrng[2], perchg),
    negs = sign(perchg),  
    rnks = scales::rescale(abs(perchg), c(1, 1000)), 
    rnks = round(rnks), 
    cols = ifelse(negs == 1, palpos[rnks], palneg[rnks]),
    ind = 1:nrow(.), 
    cattype = ifelse(cat %in% molabs, 'mo', 'an')
  )
# plot(perchg ~ ind, dat = toplo, col = cols)

unis <- with(toplo, paste(resvar, cattype)) %>% 
  unique %>% 
  sort

@@

  # add site labels if plot is first
  pall <- pbase
  set.seed(5)
  if(which(uni == unis) == 1)
    pall <- pbase + 
      geom_label_repel(
        data = statmeta, 
        aes(x = Longitude, y = Latitude, label = Site_Code), 
        label.r = unit(0, "lines"),
        box.padding = unit(2, "lines"),
        label.padding = unit(0.15, "lines"), 
        point.padding = unit(0.1, "lines"),
        force = 2, size = 2,
        fill = strp_fl
      )

@@

      nms <- names(bars)[i]
      dat <- statmeta[statmeta$Site_Code %in% nms,]
      pall <- subview(pall, bars[[i]], x = dat$Longitude, y = dat$Latitude, subwid, 0.14)

@@

# setup legend
leg <- ggplot(data.frame(x = 1:10, y = 1:10, z = seq(-5, 5, length = 10)), aes(x, y, colour = z)) + 
  geom_point() + 
  scale_colour_gradientn(colours = c(rev(colneg), colpos)) + 
  guides(colour = guide_colourbar(title = expression('% chg yr' ^-1), barheight = 5, barwidth = 0.5))
leg <- g_legend(leg)

grid.arrange(
  arrangeGrob(plo_ls[[1]], plo_ls[[3]], plo_ls[[5]], ncol = 1, top = c('Annual (1976-1995, 1996-2014)')), 
  arrangeGrob(plo_ls[[2]], plo_ls[[4]], plo_ls[[6]], ncol = 1, top = c('Seasonal (Sp, Su, Fa, Wi)')),
  leg,
  ncol = 3, 
  widths = c(1, 1, 0.4)
  )

@@

% overall trends
<<trnds>>=
data(trnds_sk)

>>>>

source('R/funcs.r')
library(WRTDStidal)
library(ggplot2)
library(dplyr)
library(gridExtra)
library(grid)
library(purrr)
library(tidyr)
library(Hmisc)
library(ggrepel)
library(ggtree)
library(forcats)
@

@@

<<trndmap, fig.cap = 'Percent change per year in nitrogen analytes for aggregations by (a) year, (b) seasons from 1976-1995, and (c) seasons from 1996-2014.  Changes are based on seasonal Kendall tests of flow-normalized results within each time period. Station names are shown in the top left panel. Values less than -9\\% change were truncated for visual clarity of the bar plots, see \\cref{tab:trnds} for actual values. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.', fig.height = 8, fig.width = 8, cache = T, out.width = '\\textwidth'>>=

# load required data
data(delt_dat)
data(delt_map)
data(trnds_seasyr)

@@

# bounding box
buffx <- 0.00035
buffy <- 0.0022
lims<- select(statmeta, -Site_Code) %>% 
  apply(., 2, function(x) range(x, na.rm = TRUE))
lims <- c(
  (1 + buffx) * lims[1, 2], 
  (1 - buffy) * lims[1, 1], 
  (1 - buffx) * lims[2, 2], 
  (1 + buffy) * lims[2, 1]
  )    
names(lims) <- c('left', 'bottom', 'right', 'top')

# need to shift locations of some stations
statmeta[statmeta$Site_Code %in% 'D26', 'Latitude'] <- 38.12
statmeta[statmeta$Site_Code %in% 'D26', 'Longitude'] <--121.55
statmeta[statmeta$Site_Code %in% 'D7', 'Latitude'] <- 38.13
statmeta[statmeta$Site_Code %in% 'D7', 'Longitude'] <--122.015 
statmeta[statmeta$Site_Code %in% 'C10', 'Latitude'] <-37.68 
statmeta[statmeta$Site_Code %in% 'C10', 'Longitude'] <--121.28

# response labels 
reslabs <- c('din an', 'nh an', 'no23 an', 'din mobef', 'nh mobef', 'no23 mobef', 'din moaft', 'nh moaft', 'no23 moaft')
reslevs <- rep(c(
  'DIN', 
  expression(paste(NH [4] ^ '+')),
  expression(paste(NO [2] ^ '-', ' / ', NO [3] ^ '2-'))
  ), 3)

@@

# bounds on limits
yrng <- c(-9, 9)

@@

# unnest data frame, assign colors by rank and sign
toplo <- select(trnds_seasyr, Site_Code, resvar, ann, cat, norm) %>% 
  rename(perchg = norm) %>% 
  left_join(., statmeta, by = 'Site_Code') %>% 
  data.frame %>% 
  mutate(
    perchg = pmax(yrng[1], perchg),
    perchg = pmin(yrng[2], perchg),
    negs = sign(perchg),  
    rnks = scales::rescale(abs(perchg), c(1, 1000)), 
    rnks = round(rnks), 
    cols = ifelse(negs == 1, palpos[rnks], palneg[rnks]),
    ind = 1:nrow(.)
    ) %>%
  unite('cattype', cat, ann, sep = '', remove = F) %>%
  mutate(
    cattype = factor(cattype),
    cattype = fct_collapse(cattype, 
      an = c('1976-1995bef', '1996-2014aft'),
      mobef = paste0(c('Spring', 'Summer', 'Fall', 'Winter'), 'bef'), 
      moaft = paste0(c('Spring', 'Summer', 'Fall', 'Winter'), 'aft') 
    )
  )

# unite response varibles and summary period combos
unis <- with(toplo, paste(resvar, cattype)) %>% 
  unique %>% 
  sort

@@

  # add site labels if plot is first
  pall <- pbase
  set.seed(5)
  if(which(uni == unis) == 1)
    pall <- pbase + 
      geom_label_repel(
        data = statmeta, 
        aes(x = Longitude, y = Latitude, label = Site_Code), 
        label.r = unit(0, "lines"),
        box.padding = unit(2, "lines"),
        label.padding = unit(0.15, "lines"), 
        point.padding = unit(0.1, "lines"),
        force = 2, size = 2,
        fill = 'grey90'
      )

@@

      nms <- names(bars)[i]
      dat <- statmeta[statmeta$Site_Code %in% nms,]
      pall <- subview(pall, bars[[i]], x = dat$Longitude, y = dat$Latitude, subwid, 0.148) # change bar heights here

@@

# setup legend
rngs <- seq(-9, 7, by = 3)
leg <- ggplot(data.frame(x = seq_along(rngs), y = seq_along(rngs), z = rngs), aes(x, y, colour = z)) + 
  geom_point() + 
  scale_colour_gradientn(colours = c(rev(colneg), colpos)) + 
  guides(colour = guide_colourbar(title = expression('% chg yr' ^-1), barheight = 0.5, barwidth = 8)) + 
  theme(legend.position = 'top')
leg <- g_legend(leg)

grid.arrange(
  leg,
  arrangeGrob(
    arrangeGrob (plo_ls[[1]], plo_ls[[4]], plo_ls[[7]], ncol = 1, top = c('(a) 1976-1995, 1996-2014')), 
    arrangeGrob(plo_ls[[3]], plo_ls[[6]], plo_ls[[9]], ncol = 1, top = c('(b) Sp, Su, Fa, Wi (1976-1995)')),
    arrangeGrob(plo_ls[[2]], plo_ls[[5]], plo_ls[[8]], ncol = 1, top = c('(c) Sp, Su, Fa, Wi (1996-2014)')),
    ncol = 3
    ),
  ncol = 1, 
  heights = c(0.1, 1)
  )

@@

% overall trends, years
<<trnds>>=
data(trnds_sk)

### 2016-12-11 sftrends_manu.tex
redid readme, deleted unused data files
fig 5 now only tot nitrogen
added seas by year to trend map

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$, Phil Trowbridge$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Mon Nov 21 10:45:35 2016 -0600
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

\caption{Percent change per year in nitrogen analytes for (a) annual and (b) seasonal aggregations.  Changes are based on seasonal Kendall tests of flow-normalized results within each time period. Station names are shown in the top left panel. Three values less than 5\% change were truncated for visual clarity of the bar plots, see \cref{tab:trnds} for actual values. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.}\label{fig:trndmap}
\end{figure}

@@

% overall trends

>>>>

\begin{singlespace}
\title{{\bf {\Large Four decades of water quality change in the upper San Francisco Estuary}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, David Senn$^2$, Phil Bresnahan$^2$, Emily Novick$^2$, James D. Hagy III$^1$,}}
  \\{\bf {\normalsize Thomas Jabusch$^2$, Phil Trowbridge$^2$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:hagy.jim@epa.gov}{hagy.jim@epa.gov}}}
  \\\\{\textit {\normalsize $^2$San Francisco Estuary Institute}}
	\\{\textit {\normalsize 4911 Central Avenue, Richmond, CA 94804}}
	\\{\textit {\normalsize Phone: 510-746-7334, Fax: 510-746-7300}}
	\\{\textit {\normalsize Emails: \href{mailto:davids@sfei.org}{davids@sfei.org}, \href{mailto:philb@sfei.org}{philb@sfei.org}, \href{mailto:emilyn@sfei.org}{emilyn@sfei.org}, \href{mailto:thomasj@sfei.org}{thomasj@sfei.org}}}
  \vspace{1in} 
  \\ Version Date:   Fri Dec 9 16:31:55 2016 -0600
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

\caption{Percent change per year in nitrogen analytes for aggregations by (a) year, (b) seasons from 1976-1995, and (c) seasons from 1996-2014.  Changes are based on seasonal Kendall tests of flow-normalized results within each time period. Station names are shown in the top left panel. Values less than -9\% change were truncated for visual clarity of the bar plots, see \cref{tab:trnds} for actual values. Months for each season are Spring: MAM, Summer: JJA, Fall: SON, Winter: DJF.}\label{fig:trndmap}
\end{figure}

@@

% overall trends, years
